\renewcommand{\hwnum}{3}

\title{Third Homework Assignment}
\date{due: 23 February 2022}
\maketitle

This homework is worth $7$ points.
\input{body/hw-directions}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\collab{n/a}
\nextprob{Getting to Know You (1 point)}

\begin{enumerate}

    \item Change the photo in D2L to be a headshot so that I can recognize who you are.
        \paragraph{Answer}
        \todo{After you are done, remove this ``TODO'' and write `Done'.}

    \item What do you want to do after you graduate? (If submitting as a group,
        state what each individual member of the group hopes to do).
        \paragraph{Answer}
        \todo{}

    \item You have been in this class for a couple of weeks now.  Why do you
        think this class is so important for computer science majors?
        \paragraph{Answer}
        \todo{}

    \item What is the ``muddiest point'' in this class so far?  (That is, what
        concept / proof / definition have you seen that you are the least
        comfortable with?)  If there are multiple ``muddy points,'' list them.
        \paragraph{Answer}
        \todo{}

    \item This HW should be submitted as a PDF.
        For extra credit, submit a typeset document. For extra extra credit,
        use LaTex.

        No response needed here. Just select any page in Gradescope
        for this one.

\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\collab{\todo{}}
\nextprob{Review and Practice Questions (1 point)}

\begin{enumerate}
    \item Is the following statement true: If $p$ is a prime number, then $p$ is an odd
        integer. Justify.\label{part:thestmt}
        \paragraph{Answer}
        \todo{}

    \item Negate the statement in Part \ref{part:thestmt}. above.
        \paragraph{Answer}
        \todo{}

    \item Is the negation in the previous part true or false? Justify.
        \paragraph{Answer}
        \todo{}

    \item Give the contrapositive of the statement in Part \ref{part:thestmt}. above.
        \paragraph{Answer}
        \todo{}

    \item Is the contrapositive that you gave in the previous part true or
        false? Justify.
        \paragraph{Answer}
        \todo{}

    \item Your friend gave you the following proof that multiples of $4$ are
        even: \emph{Let $n=100$.  Since $n=4\cdot 25$, we have that $n$ is a multiple
        of $4$.  Also, $n=2\cdot 50$; therefore, by definition of even, $n$ is
        even.}  Does this proof work?  Why or why not?
        \paragraph{Answer}
        \todo{}

    \item What is the definition of an even number?
        \paragraph{Answer}
        \todo{}

    \item Give a counter-example to the following statement: for $a,b \in \Z$,
        $$a^2=b^2 \implies a=b.$$
        \paragraph{Answer}
        \todo{}

\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\collab{\todo{}}
\nextprob{Divisibility (1 point)}

Suppose $n$ is an integer whose decimal representation ends in $5$.  Prove that
$5$ divides $n$. (Hint: recall that the decimal number $532$ can be written as
the sum $5\cdot 100 + 3 \cdot 10 + 2 \cdot 1$).

\paragraph{Answer}
\todo{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\collab{\todo{}}
\nextprob{Quotient-Remainder Theorem (1 point)}

Let $n$ be an integer.  Prove that $n$ can be written in one of the three forms:
$$
    n=3q
    \text{ or }
    n=3q+1
    \text{ or }
    n=3q+2
$$
for some integer $q$. (Hint: Use the QRT).

\paragraph{Answer}
\todo{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\collab{\todo{}}
\nextprob{Indirect Proof (1 point)}

Consider the following statement:

For every integer $n$, if $n^2$ is odd, then
$n$ is odd.

\begin{enumerate}
    \item Prove that this statement is correct.
        \paragraph{Answer}
        \todo{}

    \item Explain what proof technique you used to prove this (direct proof,
        proof by contradiction, proof by contrapositive, other).
        \paragraph{Answer}
        \todo{}

    \item What is the converse of the if/else statement, and is the converse true?
        \paragraph{Answer}
        \todo{}
\end{enumerate}
