\renewcommand{\hwnum}{4}
\setcounter{probnum}{0}

\title{Fourth Homework Assignment}
\date{due: 2 March 2022}
\maketitle

This homework is worth $7$ points.

\input{body/hw-directions}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\collab{\todo{}}
\nextprob{Review and Practice Questions (1 point)}

\begin{enumerate}
    \item Does $6$ divide $180$? Justify.
        \paragraph{Answer}
        \todo{}

    \item What is the biggest remainder possible when you divide a number by $8$?
        \paragraph{Answer}
        \todo{}

    \item Is the following statement true or false? \emph{If $n$ is an integer,
        then $n$ is a rational number}. (Briefly~justify). \label{part:newstmt}
        \paragraph{Answer}
        \todo{}

    \item What is the contrapositive of the statement in Part~\ref{part:newstmt}? Is
        it true? (Briefly~justify).
        \paragraph{Answer}
        \todo{}

    \item What is the converse of the statement in Part~\ref{part:newstmt}? Is
        it true? (Briefly justify).
        \paragraph{Answer}
        \todo{}

    \item If we were to (attempt to) prove the statement in
        Part~\ref{part:newstmt} by contradiction, what would the supposition~be?
        \paragraph{Answer}
        \todo{}

    \item Use the Fundamental Theorem of Arithmetic to write $378$ as a product
        of prime numbers.
        \paragraph{Answer}
        \todo{}

    \item Find $q,r$ from the QRT for $n=25$ and $d=4$.
        \paragraph{Answer}
        \todo{}

\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\collab{\todo{}}
\nextprob{Sequences (1 point)}

\begin{enumerate}

    \item What is the $n$-th number in the following sequence:
        $$
            0, -\frac{1}{3}, \frac{4}{9}, -\frac{1}{3}, \frac{16}{81},
            -\frac{25}{243}, \frac{36}{729}, \ldots?
        $$
        (For reference,
        the zeroth number is $0$, the $1$st number is $-\frac{1}{3}$, and the $3$rd
        number is $\frac{4}{9}$).
        \paragraph{Answer}
        \todo{}

    \item Let $a_0=0$ and let $a_k=2a_0+k$, for $k\geq 1$.
        Compute $\sum_{k=0}^3 a_{2k}$.
        \paragraph{Answer}
        \todo{}

    \item Let $b_k=\frac{k}{k+1}$, for $k\in \N$.
        Compute  $\prod_{i=0}^{10} b_{k}$.
        \paragraph{Answer}
        \todo{}

    \item Consider the following set: $\{ 1/n\}_{n=1}^{\infty} \cup \{ 0 \}$.
        Is this set well-ordered. Why or why not?
        \paragraph{Answer}
        \todo{}



\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\collab{\todo{}}
\nextprob{Proof by Induction (2 points)}

Induction is an important concept going forward in your computer science
curriculum.  Please try (or re-try) questions in the lectures and the textbook,
as two induction problems is not enough to master the concept.  Prove the
following two statements using induction.

\begin{enumerate}
    \item The number of edges in a complete graph on $n$ vertices is
        $\frac{n(n-1)}{2}$. % Section 5.3, Problem 31

        \paragraph{Answer}
        \todo{}

    \item Suppose that $c_0,c_1,c_2,\ldots$ is the sequence defined as follows:
        $c_0=2$, $c_1=2$,and, for every $k \geq 2$, $c_k:=c_{k-1}+3\cdot c_{k-2}$.
        Prove that $c_n$ is even for each integer $n \geq 0$.

        \paragraph{Answer}
        \todo{}

\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\collab{\todo{}}
\nextprob{Recursion (2 points)}

(Section 5.6, Exercise 40). A set of blocks contains blocks of heights $1$, $2$,
and $4$ centimeters.  Imagine constructing towers by piling blocks of different
heights directly on top of one another. (A tower of height $6$ cm could be
obtained using six $1$-cm blocks; three $2$-cm blocks; one $2$-cm block on the
bottom with one $4$-cm block on top; one $4$-cm block on bottom with one $2$-cm
block on top, and so forth.) Let $t_n$  be the number of ways to construct a
tower of height $n$ cm using blocks from the set. (Assume an unlimited supply of
blocks of each size.) Find a recurrence relation for $t_1, t_2, t_3, \ldots $.
Justify your answer.

\paragraph{Answer}
\todo{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\collab{\todo{}}
\nextprob{More Recurrence Relations (1 point)}

Find the closed form of the following recurrence relations:

\begin{enumerate}
    \item $T(n) = T(n-1) + 5$, with $T(0)=1$.
        \paragraph{Answer}
        \todo{}

    \item $T(n) = 2 T(n-1) + n$, with $T(0)=1$.
        \paragraph{Answer}
        \todo{}

    \item $T(n) = 2 T(n-1) + 3^n$, with $T(0)=1$.
        \paragraph{Answer}
        \todo{}

    \item (Bonus) What is the asymptotic form of the closed forms above?
        \paragraph{Answer}
        \todo{}
\end{enumerate}
