\section*{Lecture 12: Quotient-Remainder Theorem (\textsection 4.5)}

The following lemma is widely known as \emph{Euclid's Division Lemma}, but our
textbook calls it the \emph{Quotient Remainder Theorem}.  So, we will stick with
that name.  We state it here without proof, but the proof is an exercise in the
textbook (Exercise 21 of Section 4.8).

\begin{theorem}[Quotient-Remainder Theorem (QRT)]
    For all integers $n$ and $d$ such that $d > 1$,
    there exists unique integers $q$
    and $r$ such that
    $$ n = dq + r, $$
    where $0 \leq r < d$.
\end{theorem}

As computer scientists, we are familiar with this theorem.  In fact, we know
these as integer division (``div'') and modular arithmetic (``mod''):
$$q = n \text{ div } d \text{, and } r= n\mod d. $$
Different programming languages this slightly differently syntax for mod and
div (e.g., integer
division is~$n//d$
in python, but~$n/d$ in C++).

\input{body/prac-directions}

\begin{enumerate}
    \item Find $q,r$ from the QRT for $n=27$ and $d=5$.
        \practice
    \item Find $q,r$ from the QRT for $n=25$ and $d=5$.
        \practice
    \item Find $q,r$ from the QRT for $n=21$ and $d=5$.
        \practice
    \pagebreak
    \item Find $q,r$ from the QRT for $n=0$ and $d=5$.
        \practice
    \item Find $q,r$ from the QRT for $n=-8$ and $d=5$.
        \practice
    \item Find $r$ from the QRT if $n$, $d$, and $q$ are known.
        \practice
    \item Today is Wednesday, 16 February 2022.  What day of the week is
        Wednesday, 18 February 2023? (Note: 2022 is not a leap year).
        \practice
\end{enumerate}

Note that $d=2$ is a special case.  In fact, $n$ is even iff $n \mod
2 = 0$.  And, $n$ is odd iff $n \mod 2 = 1$.  The \emph{parity} of an integer
refers to whether the integer is even or odd.

\begin{theorem}[Divisibility is Transitive]
    A necessary and sufficient condition for an integer $n$ to be divisible by a
    positive integer $d$ is that $n \mod d = 0$.
\end{theorem}
\begin{proof}
    Let $n \in \Z$ and let $d \in \Z_{+}$.

    First, we will prove the forward direction: \emph{If $n$ is divisible by
    $d$, then $n \mod d = 0$.}
    \proofspace

    Next, we will prove the backward direction: \emph{If $n \mod d = 0$,
    then $n$ is divisible by $d$.}
    \proofspace

    Thus, we have shown that $n \mod d = 0$ is a necessary and sufficient
    condition for $n$ to be divisible by $d$.
\end{proof}
